---
title: "Simulation of IHPP for different intensity function"
author: "Lucas Rosas, Jeanne Marque, Clélia Desjean and Marwan Noua (4A MA)"
date: "2025/2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generative Functions

For this part we used the functions used in the TP

```{r simulPPh}
simulHPP <- function(lambda, Tmax) {
  N <- rpois(1, lambda * Tmax)
  U <- runif(N, 0, Tmax)
  return(sort(U))
}
```

```{r plot.PP}
plot.PP <- function(PP) {
  plot(c(0, PP), 0:length(PP), type = "s",
       xlab = "time t", ylab = "number of events by time t")
  points(PP, 0 * PP, type = "p", pch = 16)
  lines(PP, 0:(length(PP) - 1), type = "h", lty = 2)
}
```

```{r simulPPi}
simulPPi <- function(lambda_fct, Tmax, M) {
  res  <- NULL
  PPh  <- simulHPP(M, Tmax)
  if (length(PPh) == 0) return(NULL)
  U <- runif(length(PPh), 0, M)
  for (i in 1:length(PPh)) {
    if (U[i] < lambda_fct(PPh[i])) res <- append(res, PPh[i])
  }
  return(sort(res))
}
```

------------------------------------------------------------------------

# Inhomogeneous Poisson Processes

## Intensity functions

We define different intensity functions such that for $H_0 : \beta =1$ the intensity has a constant rate.

```{r intensities}
Weibull_I <- function(x, beta, theta) {(beta / theta) * (x / theta)^(beta - 1)}

I_sigm <- function(t, beta, theta) {1 + (beta - 1) / (1 + exp(-(t - theta)))}


I_sqrt <- function(t, beta, theta){ (1/theta) * sqrt(1 + (beta - 1) * t / theta)}
I_log <- function(t, beta, theta) {(1/theta) * (1 + (beta - 1) * log(1 + t / theta))}
```

## Simulation functions

We now define the function needed to simulate the PP for each intensity function

```{r simulations}
simulWeibull <- function(beta, theta, Tmax) {
  f     <- function(t) {Weibull_I(t, beta, theta)}
  M_val <- f(Tmax)
  return(simulPPi(f, Tmax, M_val))
}


simulLog <- function(beta, theta, Tmax) {
  f     <- function(t) {I_log(t, beta, theta)}
  M_val <- f(Tmax)
  return(simulPPi(f, Tmax, M_val))
}

simulSigm <- function(beta, theta, Tmax) {
  f     <- function(t) {I_sigm(t, beta, theta)}
  M_val <- f(Tmax)
  return(simulPPi(f, Tmax, M_val))
}

simulSqrt <- function(beta, theta, Tmax) {
  f     <- function(t) {I_sqrt(t, beta, theta)}
  M_val <- f(Tmax)
  return(simulPPi(f, Tmax, M_val))
}


```

## Simulation and plot

Let's see !

```{r plot}
Beta_val  <- 1.4
Theta_val <- 1
Tmax_val  <- 100

par(mfrow = c(1, 2))
simW <- simulWeibull(Beta_val, Theta_val, Tmax_val)
plot.PP(simW); title("Weibull Process")
curve(Weibull_I(x, Beta_val, Theta_val), 0, Tmax_val, col = "red", lwd = 2, main = "Weibull Intensity")

par(mfrow = c(1, 2))
simL <- simulLog(Beta_val, Theta_val, Tmax_val)
plot.PP(simL); title("Log Process")
curve(I_log(x, Beta_val, Theta_val), 0, Tmax_val, col = "red", lwd = 2, main = "Log Intensity")


par(mfrow = c(1, 2))
simSg <- simulSigm(Beta_val, Theta_val, Tmax_val)
plot.PP(simSg); title("Sigmoid Process")
curve(I_sigm(x, Beta_val, Theta_val), 0, Tmax_val, col = "red", lwd = 2, main = "Sigmoid Intensity")


par(mfrow = c(1, 2))
simS <- simulSqrt(Beta_val, Theta_val, Tmax_val)
plot.PP(simS); title("Sqrt Process")
curve(I_sqrt(x, Beta_val, Theta_val), 0, Tmax_val, col = "red", lwd = 2, main = "Sqrt Intensity")
```

------------------------------------------------------------------------

# Statistical Tests

We now want to implement the Crow and Barlow test presented in the paper.

We added a silent parameter to choose whether we want a text response in the style of t.test.

## Crow test

```{r testCrow}
testCrow <- function(PP, Tmax, silent = FALSE)
{
  n <- length(PP)
  
  stat_test_Obs = 2 * sum(log(Tmax / PP)) 
  
  p_value <- pchisq(stat_test_Obs, 2 * n)
  
  if (!silent) {
    print(paste("The Observed Crow Statistic is : ", stat_test_Obs))
    print(paste("The p-value : ", p_value))
    
    if (p_value < 0.05) {
      print("Conclusion: we reject H0 at level 5%, Beta is greater than 1")
    } else {
      print("Conclusion: we do not reject H0 at level 5%, Beta could be 1")
    }
  }

  return(p_value)
}
```

## Barlow test

```{r testBarlow}
testBarlow <- function(PP, Tmax, silent = FALSE)
{
  n = length(PP)
  if (n < 4) return(NA) 
  
  d = floor(n/2)
  
  stat_test_Obs = ((n - d) * PP[d]) / (d * (PP[n] - PP[d]))
  
  p_value <- 1 - pf(stat_test_Obs, 2 * d, 2 * (n - d))
  
  if (!silent) {
    print(paste("The Observed test Statistic is : ", stat_test_Obs))
    print(paste("The p-value : ", p_value))
    
    if (p_value < 0.05) {
      print("Conclusion: we reject H0 at level 5%, Beta is greater than 1")
    } else {
      print("Conclusion: we do not reject H0 at level 5%, data suggests Beta could be 1")
    }
  }

  return(p_value)
}

```

## Example test execution

```{r test_execution}
Tmax <- 100
set.seed(42)
simEx <- simulWeibull(beta = 1.5, theta = 1, Tmax = Tmax)

print("------ Barlow ------")
testBarlow(simEx, Tmax)

print("------ Crow ------")
testCrow(simEx, Tmax)
```

------------------------------------------------------------------------

Power Analysis We now want to analyze our test’s power. To do so, we will first see how the tests respond to different values of $\beta$ and then to different values of $T_{\max}$.

## 1. Power as a function of Beta

```{r power_vs_beta}
plot_puissance_beta <- function(simul_func, liste_beta, Tmax, nsimu,
                                theta = 1, titre = "") {
  puissances_C <- numeric(length(liste_beta))
  puissances_B <- numeric(length(liste_beta))

  for (j in seq_along(liste_beta)) {
    b            <- liste_beta[j]
    rejet_crow   <- 0
    rejet_barlow <- 0

    for (i in 1:nsimu) {
      donnees <- simul_func(beta = b, theta = theta, Tmax = Tmax)
      if (length(donnees) > 1) {
        p_c <- testCrow(donnees, Tmax, silent = TRUE)
        p_b <- testBarlow(donnees, Tmax, silent = TRUE)
        if (!is.na(p_c) && p_c < 0.05) rejet_crow   <- rejet_crow   + 1
        if (!is.na(p_b) &&p_b < 0.05) rejet_barlow <- rejet_barlow + 1
      }
    }
    puissances_C[j] <- rejet_crow   / nsimu
    puissances_B[j] <- rejet_barlow / nsimu
  }

  plot(NULL, xlim = range(liste_beta), ylim = c(0, 1),
       xlab = "Beta", ylab = "Power (rejection rate)",
       main = paste("Power vs Beta —", titre, "| Tmax =", Tmax))
  abline(h = 0.05, col = "darkgrey", lty = 2)
  lines(liste_beta, puissances_C, col = "blue", lwd = 2, type = "b", pch = 19)
  lines(liste_beta, puissances_B, col = "red",  lwd = 2, type = "b", pch = 17)
  legend("topright", legend = c("Crow", "Barlow"),
         col = c("blue", "red"), lwd = 2, pch = c(19, 17), bty = "n")
}
```

```{r run_power_beta, fig.width=10, fig.height=5}
betas_test <- seq(1, 1.6, by = 0.1)
nsimu      <- 20

par(mfrow = c(2, 2))
plot_puissance_beta(simulWeibull, betas_test, Tmax = 100, nsimu = nsimu,
                    theta = 1, titre = "Weibull")


plot_puissance_beta(simulLog, betas_test, Tmax = 100, nsimu = nsimu,
                    theta = 1, titre = "Log")

plot_puissance_beta(simulSqrt, betas_test, Tmax = 100, nsimu = nsimu,
                    theta = 1, titre = "Sqrt")

plot_puissance_beta(simulSigm, betas_test, Tmax = 100, nsimu = nsimu,
                    theta = 1, titre = "Sigmoid")
```

------------------------------------------------------------------------

## 2. Power as a function of Tmax (fixed Beta)

```{r power_vs_tmax}
plot_puissance_tmax <- function(simul_func, liste_Tmax, beta_fixe, nsimu,
                                theta = 1, titre = "") {
  puissances_C <- numeric(length(liste_Tmax))
  puissances_B <- numeric(length(liste_Tmax))

  for (j in seq_along(liste_Tmax)) {
    Tmax         <- liste_Tmax[j]
    rejet_crow   <- 0
    rejet_barlow <- 0

    for (i in 1:nsimu) {
      donnees <- simul_func(beta = beta_fixe, theta = theta, Tmax = Tmax)
      if (length(donnees) >= 4) {
        p_c <- testCrow(donnees, Tmax, silent = TRUE)
        p_b <- testBarlow(donnees, Tmax, silent = TRUE)
        if (!is.na(p_c) &&p_c < 0.05) rejet_crow   <- rejet_crow   + 1
        if ( !is.na(p_b) &&p_b < 0.05) rejet_barlow <- rejet_barlow + 1
      }
    }
    puissances_C[j] <- rejet_crow   / nsimu
    puissances_B[j] <- rejet_barlow / nsimu
  }

  plot(NULL, xlim = range(liste_Tmax), ylim = c(0, 1),
       xlab = "Tmax (observation window)", ylab = "Power (rejection rate)",
       main = paste("Power vs Tmax —", titre, "| Beta =", beta_fixe))
  abline(h = 0.05, col = "darkgrey", lty = 2)
  lines(liste_Tmax, puissances_C, col = "blue", lwd = 2, type = "b", pch = 19)
  lines(liste_Tmax, puissances_B, col = "red",  lwd = 2, type = "b", pch = 17)
  legend("topright", legend = c("Crow", "Barlow"),
         col = c("blue", "red"), lwd = 2, pch = c(19, 17), bty = "n")
}
```

```{r run_power_tmax, fig.width=10, fig.height=5}
liste_Tmax <- seq(10, 200, by = 20)
beta_fixe  <- 1.3
nsimu      <- 200

par(mfrow = c(2, 2))
plot_puissance_tmax(simulWeibull, liste_Tmax, beta_fixe = beta_fixe,
                    nsimu = nsimu, theta = 1, titre = "Weibull")

plot_puissance_tmax(simulLog, liste_Tmax, beta_fixe = beta_fixe,
                    nsimu = nsimu, theta = 1, titre = "Log")

plot_puissance_tmax(simulSigm, liste_Tmax, beta_fixe = beta_fixe,
                    nsimu = nsimu, theta = 1, titre = "Sigmoid")

plot_puissance_tmax(simulSqrt, liste_Tmax, beta_fixe = beta_fixe,
                    nsimu = nsimu, theta = 1, titre = "Sqrt")

```

------------------------------------------------------------------------

# The sigmoid Function

Crow and Barlow’s tests are designed to assess whether events accumulate progressively faster over the entire observation interval. When the underlying intensity follows a sigmoid shape, the process typically exhibits three phases: an initial flat period, a rapid increase, and a final flat period.

The location of the transition point, denoted by \\theta, strongly influences the behavior of the tests:

-   Small $\theta$ (early jump): Events are concentrated at the beginning of the interval and become sparse later. This pattern may resemble a decreasing intensity, which can lead the tests to incorrectly reject in the wrong direction.

-   $\theta = T_{\max}/2$ (midpoint jump): The signal is mixed, with no clear dominance of either early or late accumulation, making the tests less decisive.

-   Large $\theta$ (late jump): The process remains nearly flat for most of the interval, providing little evidence of any trend. As a result, the tests have low power.

Overall, this behavior shows that Crow and Barlow’s tests implicitly rely on the assumption of a globally and smoothly increasing intensity. They are therefore sensitive to the shape of the alternative and may perform poorly when the true intensity deviates from this assumption.

```{r, sigmoidpower}


par(mfrow = c(2, 2))
plot_puissance_beta(simulSigm, betas_test, Tmax = 100, nsimu = nsimu,
                    theta = 1, titre = "Sigmoid, Theta = 1")

plot_puissance_beta(simulSigm, betas_test, Tmax = 100, nsimu = nsimu,
                    theta = 50, titre = "Sigmoid, Theta = 50")

plot_puissance_beta(simulSigm, betas_test, Tmax = 100, nsimu = nsimu,
                    theta = 100, titre = "Sigmoid, Theta = 100")



```